SNP Analysis

### 2018-10-12 AllSamplesModel.Rmd
Using new package

ggsave('plots/prediction_nohgd.png', plot=hgdloo, width=5, height=5, units='in', dpi=300)

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
  geom_point() + 
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
nohgdcalib

ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

@@

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[hgd.samples$Samplename,]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  hgd.samples$Prediction = pred

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

#hist(hgd.samples$Prediction)

rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r echo=F, fig.height=4, fig.width=4}
#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

@@

ft.fun<-function(NP,P) {
  #f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  f = chisq.test(c(NP,P))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
pred.confidence = qt %>% as.data.frame.matrix %>% rowwise() %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .3)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

@@

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
#pred.confidence$Risk[4:7] = 'Moderate'

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
  plot.theme + theme(legend.position = 'none')
pcal

@@

ggplot(back, aes(Status, OR, color=OR)) + 
  geom_boxplot(outlier.colour=NA) + geom_jitter(width=0.3, aes(shape=Pathology)) + 
  scale_color_gradientn(colors = myPal,  name='') + plot.theme

@@

ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
```

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
back$ogj = back$nl
back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
back$ogj[which(back$nl >= 7)] = 3

@@

p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

@@

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)

p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'' )
grid.arrange(p2$grob)

@@

get.w.scale <- function(b) {
  n = sum((b %>% group_by(Patient) %>% summarise(unique(n.endos)))[,2])
  #round(n*0.5)+1
  n
}

@@

x = sapply(subset(sum.patient.data, Status == 'NP')$Patient, function(i) {
  b = subset(back, Patient == i)
  get.h.scale(b, 9)
})
sort(x)

@@

plots = list(); widths = c()
for (i in subset(sum.patient.data, Status == 'P')$Patient) {
  b = subset(back, Patient == i)
  lim = c(1:7)
  if (b$max.block[1] < 6) lim = c(1:5)
  if (b$max.block[1] > 2 & b$max.block[1] < 4) lim = c(1:3)
  if (b$max.block[1] <= 2) lim = c(1:2)

@@

for (i in subset(sum.patient.data, Status == 'NP')$Patient) {
  b = subset(back, Patient == i)
  lim = c(1:9)
  if (b$max.block[1] < 6) lim = c(1:5)
  if (b$max.block[1] < 4) lim = c(1:3)
  if (b$max.block[1] < 3) lim = c(1:2)

@@

```{r echo=F, eval=T}
dirs = c('~/Data/Ellie/Analysis/downsampled5G_BEAdjacent',
         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/') #,'~/Data/Ellie/Analysis/downsampled5G_NDBE')

@@

predfiles = sapply(dirs, function(x) {
  list.files(x, 'predictions_all_exAHM0320.Rdata', full.names = T)
})

predObj = list()
for (i in 1:length(predfiles)) {
  if (length(predfiles[[i]]) > 0) {
      load(predfiles[[i]], verbose=T)
      dspred$Type = basename(names(predfiles[i]))
      predObj[[i]] = dspred
      rm(dspred)
  } else {
    files = list.files(names(predfiles[i]), pattern='raw', full.names=T)
    sampleNames = sub('_raw.txt', '', basename(files))
    dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
    for (n in 1:length(files)) {
      print(n)
      print(sampleNames[n])

      rs = read.table(files[n], sep='\t',header=T)
      if (ncol(rs) < 4)
        stop("File needs to contain 4 columns: chr, start, end, sample_value")

      segs <- tile.segmented.data(rs, size=5e6, chr.info=chr.lengths)
      segM = as.matrix(segs[,-c(1:3)])
      rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
      segM = t(segM)
      for (j in 1:ncol(segM)) 
        segM[,j] = unit.var(segM[,j], z.mean[j], z.sd[j])

      arms <- tile.segmented.data(rs, size='arms', chr.info=chr.lengths)
      armsM = as.matrix(arms[,-c(1:3)])
      rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
      armsM = t(armsM)
      for (j in 1:ncol(armsM)) 
        armsM[,j] = unit.var(armsM[,j], z.arms.mean[j], z.arms.sd[j])

      nrow(armsM) == nrow(segM)

      cx.score = score.cx(segM,1)

      mergedDf = subtract.arms(segM, armsM)
      mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
      dim(mergedDf)

      # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
      sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                                  dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
      for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

      preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
      RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')

      dspred[n,] = c(preds[,1],RR[,1])
      print(dspred[1:n,])
  }

    dspred$Type = basename(names(predfiles[i]))
    predObj[[i]] = dspred

    save(dspred, file=paste(names(predfiles[i]), 'predictions_all_exAHM0320.Rdata', sep='/'))
    rm(dspred)
  }

  dspred = do.call(rbind, predObj)
}

@@

dspred$Type[grep('normal', dspred$Type)] = 'Normal blood'
dspred = subset(dspred, !grepl('normal', Type))
dspred$Type[grep('BEAdjacent', dspred$Type)] = 'Tumor adjacent BE'

@@

#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
colnames(prg) = colnames(dspred)
dspred = rbind(dspred,prg)
dspred$Type = factor(dspred$Type, levels=c('Normal blood','Non-Progressor','Progressor','Tumor adjacent BE'), ordered = T)

colors = myPal[seq(2, length(myPal), round(length(myPal)/4))]
pr = ggplot(dspred, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2, aes(color=RR)) + 
  scale_color_gradientn(colors = myPal,  name='') + 
  geom_boxplot(outlier.colour = NA) + 
  scale_fill_manual(values = colors,  name='') + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45,hjust=1)) + plot.theme 
pr
ggsave('plots/downsampled_predictions_supplement.png', plot=pr, width=4, height=4.38,dpi=300, scale=1.1)

m = subset(dspred, Type != 'Normal blood')
m$Type = droplevels(m$Type)
pr = ggplot(m, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2, aes(color=RR)) + 
  scale_color_gradientn(colors = myPal,  name='') + 
  geom_boxplot(outlier.colour = NA) + 
  scale_fill_manual(values = riskCols,  name='') + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none') + plot.theme 
pr

@@

#predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')

@@

penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))

# Based on the penult pair
round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)

@@

rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )

@@

risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

@@

## Per year risk
## Evaluate over 5 years? How?
pyr = c()
for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

>>>>

ggsave('plots/prediction_nohgd.png', plot=hgdloo, width=5, height=5, units='in', dpi=300)
# pg.sampNOHGD = subset(pg.sampNOHGD, !is.na(Prediction))
# cuts = seq(0,1,0.1) 
# pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
# 
# 
# noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
# noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
# noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
# 
# ft.fun<-function(NP,P) {
#   f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
#   cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
# }
# 
# noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
# 
# noHGD.pred.confidence[c('r1','r2')] = NA
# for (i in 1:(length(cuts)-1)) 
#   noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
# 
# noHGD.pred.confidence$Risk = 'Moderate'
# noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
# noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
# 
# noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
#     data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
#                ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
# 
# noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
# 
# nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
#   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
#   scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
#   geom_point() + 
#   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
# nohgdcalib
# 
# ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

@@

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')

  hgd.samples[which(hgd.samples$Samplename %in% rownames(dysplasia.df)), 'Prediction'] = pred

  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

#hist(hgd.samples$Prediction)
rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r echo=F, fig.height=4, fig.width=4}
#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% dplyr::summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

@@

ft.fun<-function(NP,P) {
  #f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  f = chisq.test(c(NP,P))
  cbind.data.frame('p.value'= signif(f$p.value,4))
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

excluded = subset(back, is.na(Prediction))
back = subset(back, !is.na(Prediction))

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
pred.confidence = qt %>% as.data.frame.matrix %>% rowwise() %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(round(p.value,2) <= 0.05, '*', '') )

@@

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, round(p.value,2) <= 0.05 & perc < .3)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, round(p.value,2) <= 0.05 & perc > .5)) ] = 'High'

@@

annotation.table = dplyr::arrange(pred.confidence %>% dplyr::group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcfit = lm(perc~mn, data=pred.confidence)

rsid = rstandard(pcfit)
pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'

pcal = ggplot(pred.confidence, aes(mn, perc)) + geom_smooth(method='lm',formula=y~x, color='grey39', linetype='dashed', size=0.5,fullrange=T) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + 
  # geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='P:NP', title='') +
  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) +   coord_cartesian(xlim=c(0,1), ylim=c(0,1)) +
  scale_x_continuous(expand=c(0,0),limits=c(-0.5,1.5), breaks=cuts, labels=cuts) +  scale_y_continuous(expand=c(0,0),limits=c(-0.5,1.5), breaks=cuts, labels=cuts) +
  plot.theme + theme(legend.position = 'none')
pcal

@@

@@

#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
```

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=7}
back$ogj = back$nl
back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
back$ogj[which(back$nl >= 7)] = 3

@@

p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point(aes(shape=late), show.legend = F, size=2) + geom_line() + 
  theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

@@

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)
```

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'' )
grid.arrange(p2$grob)

@@

get.w.scale <- function(b) {
  n = sum((b %>% group_by(Patient) %>% dplyr::summarise(unique(n.endos)))[,2])
  #round(n*0.5)+1
  n
}

@@

x = sapply(subset(sum.patient.data, Status == 'NP')$Patient, function(i) {
  b = subset(back, Patient == i)
  get.h.scale(b, 9)
})
#sort(x)

@@

plots = list(); widths = c()
for (i in subset(sum.patient.data, Status == 'P')$Patient) {
  b = subset(back, Patient == i)
  if (nrow(b) <= 0) next
  lim = c(1:7)
  if (b$max.block[1] < 6) lim = c(1:5)
  if (b$max.block[1] > 2 & b$max.block[1] < 4) lim = c(1:3)
  if (b$max.block[1] <= 2) lim = c(1:2)

@@

for (i in subset(sum.patient.data, Status == 'NP')$Patient) {
  b = subset(back, Patient == i)
  if (nrow(b) <= 0) next

  lim = c(1:9)
  if (b$max.block[1] < 6) lim = c(1:5)
  if (b$max.block[1] < 4) lim = c(1:3)
  if (b$max.block[1] < 3) lim = c(1:2)

@@

```{r echo=F, eval=T}

@@

down.dir = '~/Data/Ellie/Analysis/downsampled'
dsids = read_xlsx(paste(data,'Analysis/downsampled_ids.xlsx',sep='/'))
load(list.files(down.dir, 'Rdata', full.names = T), verbose=T)

dspred = merge(dspred, dsids, by.x='row.names',by.y='Illumina ID')
rownames(dspred) = dspred$Row.names
dspred$Row.names = NULL
head(dspred)

@@

@@

#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
colnames(prg) = colnames(dspred)[c(1:2,ncol(dspred))]

dspred = rbind(dspred[colnames(prg)],prg)

dspred= subset(dspred, Type != 'LGD') # These are pre-progressors most likely

dspred$Type = factor(dspred$Type, levels=c('Normal', 'NDBE','Non-Progressor','Progressor','Tumor Adjacent BE'), ordered = T)

colors = myPal[seq(2, length(myPal), round(length(myPal)/length(levels(dspred$Type))))]
pr = ggplot(dspred, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2, aes(color=RR)) + 
  scale_color_gradientn(colors = myPal,  name='') + 
  geom_boxplot(outlier.colour = NA) + 
  scale_fill_manual(values = colors,  name='') + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45,hjust=1)) + plot.theme 
pr
ggsave('plots/downsampled_predictions_supplement.png', plot=pr, width=4, height=4.38,dpi=300, scale=1.1)

m = subset(dspred, !Type %in% c('Normal','LGD','NDBE'))
m$Type = droplevels(m$Type)

colors = myPal[seq(2, length(myPal), round(length(myPal)/length(levels(m$Type))))]
pr = ggplot(m, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2, aes(color=RR)) + 
  scale_color_gradientn(colors = colors,  name='') + 
  geom_boxplot(outlier.colour = NA) + 
  scale_fill_manual(values = riskCols,  name='') + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none') + plot.theme 
pr

@@

#predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')

@@

# Based on the penult pair
penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)

@@

rrPath = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )

@@

risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

@@

## Per year risk
## Evaluate over 5 years? How?
pyr = c()
for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

### 2018-10-12 snp/SNP_Prediction_Analysis.Rmd
Using new package

ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.005) + 
  geom_hline(yintercept = x) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
  geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)

@@

table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)

table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)

@@

```{r regression, echo=F, message=F, eval=T}
folds = 10; splits = 5
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

>>>>

ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.01) + 
  geom_hline(yintercept = anno$y[1]) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
  geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)

@@

#table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)

#table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)

@@

```{r regression, echo=F, message=F, eval=T}
folds = 10; splits = 5
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP/no-low/'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))
